Narrative

### 2023-05-17 manuscript/methods.qmd
added model to methods

All code and data for this project are available at (OSF LINK). For our simulations and analyses we used R (REF) and we used the R package 'metafor' (REF) to perform meta-analyses and estimate heterogeneity, For parallel computing we took advantage of the R packages 'parallel' (REF) and 'parabar' (REF).

@@

We use correlations as our effect size type in these simulations because correcting them for unreliability is straightforward and because they are the effect size type most commonly corrected in practice. However, unreliability attenuates all types of effect sizes and all can be corrected, either directly (standardized mean differences; Wiernick & Dahlke, 2020) or by first transforming them into correlations (with associated assumptions).

@@

We perform simulations both on product-moment correlations and Fisher z transformed correlations. For our purpose of estimating heterogeneity, Fisher z correlations have several advantages: 1) their sampling variance is not dependent on the size of the correlation, 2) they are not bounded to {-1, 1}. However, they also have two disadvantages: 1) that the heterogeneity estimate is measure on the Fisher z scale, which makes it hard to interpret. A workaround is to compute a prediction interval and convert this interval onto the product-moment scale (REF) for an indirect estimate of the heterogeneity. Second, adding measurement error to the Fisher z correlations requires first transforming them into product-moment correlations, adding measurement error, and then backtransforming to Fisher z for meta-analysis.

@@

@@

To add measurement error we first sampled a reliability $R_i$ for each study from a truncated ($\{0, 1\}$) normal distribution $Ntrunc(\bar{R}, \delta)$, given some average reliability across studies $\bar{R}$ and standard deviation in reliability $\delta$. We assumed that both the dependent and independent variable within a study were measured with the same reliability such that $R_i = R_{iXX'} = R_{iYY'}$. We then computed observed effect sizes $r_i$ for each study given the attenuation formula $r_{i} = \gamma_i \times \sqrt{R_{XX'}} \times \sqrt{R_{YY'}}$. Because we assumed equal reliability in $X$ and $Y$ this simplifies to $r_i = \gamma_i \times R_i$.   Fisher's $z$ we transformed to product-moment correlations before adding measurement error and then back-transformed into Fisher's $z$ before the next step. At this point we have $k$ effect sizes $r_i$ in Pearson's $r$ or Fisher's $z$ with sampling error and measurement error, ready for meta-analysis.

We meta-analyzed the $k$ effect sizes using a random effects model with inverse variance weights and the Restricted Maximum Likelihood (REML) estimator of heterogeneity. We choose the REML estimator because it is a heterogeneity estimator that performs well and is widely implemented (e.g., REF, REF2), but also show our results using the classic DerSimonian-Laird (DL) estimator in supplement XX. We used inverse variance weights due to their widespread use, but as we used a fixed sample size across studies within a meta-analysis (see next section) the weights are of little importance. As a sensitivity analysis, we also implemented the Hunter and Schmidt (2015) model which uses a different heterogeneity estimator and found qualitatively similar results (supplement XX). The procedure described in this section was replicated 10,000 times for each combination of parameter values.

>>>>

All code and data for this project are available at (OSF LINK). For our simulations and analyses we used R (REF) and we used the R package 'metafor' (REF) to perform meta-analyses and estimate heterogeneity. For parallel computing we took advantage of the R packages 'parallel' (REF) and 'parabar' (REF).

@@

We use correlations as our effect size type in these simulations because correcting them for unreliability is straightforward and because they are the effect size type most commonly corrected in practice. {see also argument in Field that all effect size types are pretty similar} However, unreliability attenuates all types of effect sizes and all can be corrected, either directly (standardized mean differences; Wiernick & Dahlke, 2020) or by first transforming them into correlations (with associated assumptions).

@@

We perform simulations both on product-moment correlations and Fisher z transformed correlations. For our purpose of estimating heterogeneity, Fisher z correlations have two advantages: 1) their sampling variance is not dependent on the size of the correlation, 2) they are not bounded to {-1, 1}. However, they also have two disadvantages: 1) that the heterogeneity estimate is measured on the Fisher z scale, which makes it hard to interpret. A workaround is to compute a prediction interval and convert this interval onto the product-moment scale (REF) for an indirect estimate of the heterogeneity. Second, adding measurement error to the Fisher z correlations requires first transforming them into product-moment correlations, adding measurement error, and then backtransforming to Fisher z for meta-analysis.

@@

## Meta-analytic model

We apply two meta-analytic models, the Hedges and Vevea (1998) random-effects  model and the Hunter and Schmidt (2015) 'bare-bones' random-effects model. We implement these two models as a sensitivity analysis rather than because of an interest in comparing the two in detail. The Hedges and Vevea (HV) model uses inverse variance weight computed as $w_i = 1 / (V_i + T^2)$ (e.g., Borenstein et al., 2010) where $V_i$ is the sampling variance of study $i$ and $T^2$ is the estimated between-studies variance. The Hunter and Schmidt (HS) model weighs studies by sample size $w_i = N_i$. The differences between these weights are relatively minor (Borenstein et al., 2010; Hunter and Schmidt, 2015, p. XX). Although the HV model is typically applied to Fisher z transformed correlations (e.g., Borenstein et al., p. 41-43), we apply it to both transformed and untransformed correlations for the reasons described in the previous section. The HS model typically includes correcting for unreliability or other measurement artifacts. However, we implement it without corrections ('bare-bones'; Hunter & Schmidt, 2015) because this allows us to examine the bias in heterogeneity estimates compared to a perfectly implemented correction of unreliability.

Estimating heterogeneity in both the HV and bare-bones HS model essentially consists of comparing the observed variance in effect sizes with what would be expected from sampling error alone and then ascribing any excess observed variance to heterogeneity (Borenstein et al., 2010, Box 1; Hunter & Schmidt, 2015, p. XX). However, the process for doing so differs somewhat between methods. We apply the Restricted Maximum Likelihood (REML) estimator of heterogeneity when estimating heterogeneity in the Hedges and Vevea (1998) model, because it is a heterogeneity estimator that performs well and is widely implemented (e.g., Viechtbauer, 2005; REF2), and is also the default estimator in the R package metafor. Describing this REML estimation in detail is beyond the scope of this paper and we refer interested readers to the description in Viechtbauer (2005) {this is also what Brannick et al do}. The Hunter and Schmidt (2015) model estimates heterogeneity $\tau^2 = \sigma^2_r - \sigma^2_e$ where  $\sigma^2_r$ is the observed variance across correlations and $\sigma^2_e$ is the average sampling error variance across studies, computed as $ \frac{\Sigma[N_i \sigma^2_{ei}]}{\Sigma N_i}$ (p. 87, 2nd edition) and $\sigma^2_{ei}$ is the usual sampling variance for the Pearson correlation but with $\rho$ replaced with the sample size weighted average correlation $\bar{r} = \frac{\Sigma n_i r_i}{\Sigma n_i}$ instead of $r_i$ as is common. A generalized version of the HS estimator for effect sizes other than correlations can be found in Viechtbauer (2005; 2015) (see ?rma).

@@

To add measurement error we first sampled a reliability $R_i$ for each study from a truncated ($\{0, 1\}$) normal distribution $Ntrunc(\bar{R}, \delta)$, given some average reliability across studies $\bar{R}$ and standard deviation in reliability $\delta$. We assumed that both the dependent and independent variable within a study were measured with the same reliability such that $R_i = R_{iXX'} = R_{iYY'}$. We then computed observed effect sizes $r_i$ for each study given the attenuation formula $r_{i} = \gamma_i \times \sqrt{R_{XX'}} \times \sqrt{R_{YY'}}$. Because we assumed equal reliability in $X$ and $Y$ this simplifies to $r_i = \gamma_i \times R_i$.   Fisher's $z$ we transformed to product-moment correlations before adding measurement error and then back-transformed into Fisher's $z$ before the next step. At this point we have $k$ effect sizes $r_i$ in Pearson's $r$ or Fisher's $z$ with sampling error and measurement error. We then applied either the HV meta-analytic model or the HS model, as described in the section 'the meta-analytic model'. The procedure described in the current section was replicated 10,000 times for each combination of parameter values.
